A studyforrest extension, an annotation of spoken language in the German dubbed movie ``Forrest Gump'' and its audio-description

### 2020-04-23 paper/main.tex
overall revision (results & tables), add discussion

% For more detailed article preparation guidelines, please see:
% http://f1000research.com/author-guidelines
\documentclass[10pt,a4paper,onecolumn]{article}
\usepackage{f1000_styles}
\geometry{right=4cm}
% enable the 'endfloat' package to sort all tables and figures to the
% end of the document without having to edit the manuscript
%\usepackage{endfloat}
\usepackage{booktabs}
\usepackage[
	colorlinks=true,
	urlcolor=blue,
	linkcolor=green
]{hyperref}
%% Default: numerical citations
\usepackage[numbers]{natbib}

@@

\author[1, 2]{Christian~O.~Häusler \\ email \href{mailto:c.haeusler@fz-juelich.de}{c.haeusler@fz-juelich.de} }
\author[1, 2]{Michael~Hanke}

@@

\thispagestyle{fancy}
\begin{abstract}
% up to 300 words
Here we present an annotation of speech in the audio-visual movie ``Forrest Gump'' and its audio-description (i.e. the audio-only variant), as an addition to a large public functional brain imaging dataset (\url{http://studyforrest.org}). The annotation provides information about the exact timing of each of the \aSentencesAll\ spoken sentences, \aWordsAll\ words (including \aPosNonspeechAll\ non-speech vocalizations), \aPhonesAll\ phonemes and its corresponding speaker.
Additionally for every word, we provide lemmatization, two different part-of-speech taggings (simple and detailed), syntactic dependencis within each sentence, and a semantic analysis in terms of word embedding (word2vector). This annotation enables further studies of language perception and semantic processing under conditions of real-life complexity using the studyforrest dataset.

Wort zur Validierungsstudie?
\end{abstract}

@@

\section*{Introduction}
\todo[inline]{Bitte am Ende (?) der Einleitung etwas zu "reproducible research" \& "datalad dataset" bringen}
Cognitive neuroimaging research is moving towards studying brain behavior
under conditions of real-life-like complexity. Motion pictures \citep{hasson2008neurocinematics} and continuous narratives \citep{honey2012not, lerner2011topographic} are increasingly utilized as so called naturalistic stimuli.

Because naturalistic stimuli are multi-dimensional and their temporal structure
is usually unknown, data-driven methods like intersubject correlation (ISC)\citep{hasson2004intersubject} or independent component analysis
(ICA)\citep{bartels2004chronoarchitecture} are often used to analyze
corresponding fMRI data.
Nevertheless, model-driven methods like a general linear model (GLM) are necessary to not just explain how but also when (and why) a brain region is responding to a stimulus \citep{hamilton2018revolution}.
Studies using GLMs based on annotations of a stimulus' temporal structure have
elucidated how the brain responds to visual features of a movie
\citep{bartels2004mapping} or speech-related features of a narrative
\citep{rocca2019language}.

\todo[inline]{In der Tabelle ausführlicher beschrieben als hier (ist okay so, denke ich): ``sentences'' sind hier nicht vollständige Sätze im linguistischen Sinne (Subjekt, Prädikat), sondern auch Ausrufe ("Forrest!") oder mitunter isolierte non-speech "Sätze" ("oouh")}
For this publication, we annotated the onsets and offsets, and content of sentences, single words, and phonemes (cf. table \ref{tab:overview} for an overview of the annotation's content) spoken in the audio-visual movie ``Forrest Gump'' \citep{ForrestGumpMovie} and its audio-description (i.e. the audio-only variant of the movie) for visually impaired persons \citep{ForrestGumpDVD}.
fMRI data of participants watching the audio-visual movie \citep{hanke2016simultaneous} and listening to the audio-description \citep{hanke2014audiomovie} are the core data of the publicly available \textit{studyforrest} dataset (\href{www.studyforrest.org}{studyforrest.org}).
Additional fMRI data of this dataset include the perception of musical genres \citep{hanke2015musical}, and retinotopic mappings and localization of higher visual areas of the same participants \citep{sengupta2016extension}.

@@

\subsection*{Annotation procedure}
\todo[inline]{wo kommen die "preliminaries transcriptions" etc.? Habe die von Medusa; die "most basic" Datei, die das Dataset enthält ist wie abgesprochen der Merge aus Dialog (ungetrennte Sentences), Narrator \& Non-Speech-Annotation = \texttt{annotation/speech-vocal.csv}}
Preliminary, orthographic transcriptions of dialogues (\texttt{german\_dialog.csv}), non-speech vocalizations (e.g. laughing or groaning; \texttt{non\_speech\_vocalization.json}), and the audio-description's narrator
(\texttt{german\_audio\_description.csv}) were merged (\texttt{speech-vocal.csv}) and converted to Praat's \citep{boersma2019praat} TextGrid format  (\texttt{speech-vocal.TextGrid}).

% \href{www.fon.hum.uva.nl/praat}{fon.hum.uva.nl/praat}.
The merged transcription containing one onset and offset for usually a couple of sentences was edited in Praat v6.0.22 for manual validation against the original audio.
The following steps were done in several passes to iteratively improve the quality of the data:
Approximate temporal onsets and offsets were corrected, and intervals containing several sentences were split into intervals containing only one sentence.
When two or more persons were speaking simultaneously the less dominant voice was dropped.
Low volume background speech (especially occurring during music or continuous environmental noise) or low volume non-speech vocalizations that are difficult to recognize by the audience were dropped, too.
\todo[inline]{what is a regular audience? without neurological disorders ;-); Simpel: eine audience, die den Film nicht 45x auf voller Lautstärke über Kopfhörer mit Transkription danebenliegend gehört hat}

% Alignment preparations
% http://prosodylab.org
% https://raw.githubusercontent.com/prosodylab/prosodylab.dictionaries/master/de.dict
We used the Montreal Forced Aligner v1.0.1 \citep{mcauliffe2017montreal} to
algorithmically identify the exact onset of offset of the sentences, and the onset and offset of each word and phoneme embedded in the transcribed sentences.
To enable the aligner to look up the phonemes embedded within each word, we chose a German pronunciation dictionary provided by Prosodylab \citep {gorman2011prosodylab} that uses the Prosodylab PhoneSet to describe the pronounciation of phonemes.
The dictionary was manually updated with German words that occur in the stimuli but were missing in the dictionary to improve the detection rate of the automatic alignment.
% \href{http://mlmlab.org/mfa/dictionaries/english.dict}{mlmlab.org/mfa/dictionaries/english.dict}
The pronunciation of English words and phonemes occurring in the otherwise German audio track was taken from an English pronunciation dictionary (following the ARPAbet PhoneSet) provided by the Montreal Language Modeling lab (\href{http://mlmlab.org/mfa/dictionaries/english.dict}{mlmlab.org}).

% Alignment
\todo[inline]{Nicht erwähnt: Wenn die Korrektur der Phoneme unmöglich war, habe ich einfach die Event-Borders zwischen den Phonemen herausgenommen. Wortbeginn und Wortende stimmen dann mit einem neuen, "künstlich-großen Phoneme" überein.
Das "Groß-Phonem" enthält dann ohne Grenzen zwischen den tatsächlichen Phonemen des Wortes alle Phoneme des Wortes ("es ist kompliziert zu erklären, aber im Grunde ganz einfach). kommt jedoch so selten vor, ist im Grunde "noise" und ist in der Anno sofort als solches zu erkennen.}
The audio track was converted from FLAC to WAV via FFmpeg v4.1.4 \citep{ffmpeg} to meet the aligner's requirements.
This WAV file, the transcription of sentences, and the updated dictionary were
submitted to the aligner that first trained an acoustic model on the data and
then performed the alignment.
The resulting .TextGrid file that contains onsets and offsets of words and phonemes was corrected manually and iteratively in several passes:
In a first step, onsets and offsets words and phonemes on which the automatic alignment performed moderately were corrected in Praat.
Some low volume sentences that are spoken in continuously noisy settings (pursuit, battle, hurricane, rally, disco) were removed due to poor overall alignment performance.
In a second step, the complete sentences of the orthographic transcription were copied into the annotation.
In a third step, the name of the speaker was added for every sentence (cf. table \ref{tab:speakers} for the most often occurring speakers).
During every step results from previous steps were checked again for errors.

\begin{table*}[tbp]
\caption{Overview of the annotation's content for the whole stimulus and its individual segments used during fMRI scanning. The label ``Sentences'' comprises complete grammatical sentences (i.e. comprising subject predicate) which are additionally marked with a full stop at the end.
The label ``Sentences'' also comprises questions (``Do you want a chocolate?'', exlamations (``Run away!), or non-speech vocalizations in quick succession (``ha, ha, ha''), or in isolation (e.g. ``Forrest?'', ``Forrest!'', ``ha'') at time points when speakers switch rapidly. The label ``Words'' comprises each word or non-speech vocalization (N=\aPosNonspeechAll) in isolation.}
\label{tab:overview}
\begin{tabular}{llllllllll}
\toprule
\textbf{label} & \textbf{all} & \textbf{1} & \textbf{2} & \textbf{3} & \textbf{4} & \textbf{5} & \textbf{6} & \textbf{7} & \textbf{8}\\
\midrule
sentences  & \aSentencesAll & \aSentencesI & \aSentencesII & \aSentencesIII & \aSentencesIV & \aSentencesV & \aSentencesVI & \aSentencesVII & \aSentencesVIII \tabularnewline
words  & \aWordsAll & \aWordsI & \aWordsII & \aWordsIII & \aWordsIV & \aWordsV & \aWordsVI & \aWordsVII & \aWordsVIII \tabularnewline
phonemes  & \aPhonesAll & \aPhonesI & \aPhonesII & \aPhonesIII & \aPhonesIV & \aPhonesV & \aPhonesVI & \aPhonesVII & \aPhonesVIII \tabularnewline
\bottomrule
\end{tabular}
\end{table*}

\begin{table*}[btp]
\caption{Sentences spoken by the ten most often occurring speakers sorted alphabetically. The narrator only occurrs in the audio-description of the movie. Names are mostly identical to the names used in \citep{labs2015portrayed}. Overall 97 persons were identified.}
\label{tab:speakers}
\begin{tabular}{llllllllll}
\toprule
\textbf{name} & \textbf{all} & \textbf{1} & \textbf{2} & \textbf{3} & \textbf{4} & \textbf{5} & \textbf{6} & \textbf{7} & \textbf{8}\\
\midrule
Bubba  & \aBubbaAll & \aBubbaI & \aBubbaII & \aBubbaIII & \aBubbaIV & \aBubbaV & \aBubbaVI & \aBubbaVII & \aBubbaVIII \tabularnewline
Forrest  & \aForrestAll & \aForrestI & \aForrestII & \aForrestIII & \aForrestIV & \aForrestV & \aForrestVI & \aForrestVII & \aForrestVIII \tabularnewline
Forrest (child)  & \aForrestchildAll & \aForrestchildI & \aForrestchildII & \aForrestchildIII & \aForrestchildIV & \aForrestchildV & \aForrestchildVI & \aForrestchildVII & \aForrestchildVIII \tabularnewline
Forrest (v.o.)  & \aForrestvoAll & \aForrestvoI & \aForrestvoII & \aForrestvoIII & \aForrestvoIV & \aForrestvoV & \aForrestvoVI & \aForrestvoVII & \aForrestvoVIII \tabularnewline
Hancock  & \aHancockAll & \aHancockI & \aHancockII & \aHancockIII & \aHancockIV & \aHancockV & \aHancockVI & \aHancockVII & \aHancockVIII \tabularnewline
Jenny  & \aJennyAll & \aJennyI & \aJennyII & \aJennyIII & \aJennyIV & \aJennyV & \aJennyVI & \aJennyVII & \aJennyVIII \tabularnewline
Jenny (child)  & \aJennychildAll & \aJennychildI & \aJennychildII & \aJennychildIII & \aJennychildIV & \aJennychildV & \aJennychildVI & \aJennychildVII & \aJennychildVIII \tabularnewline
Lt. Dan  & \aLtdanAll & \aLtdanI & \aLtdanII & \aLtdanIII & \aLtdanIV & \aLtdanV & \aLtdanVI & \aLtdanVII & \aLtdanVIII \tabularnewline
Mrs. Gump  & \aMrsgumpAll & \aMrsgumpI & \aMrsgumpII & \aMrsgumpIII & \aMrsgumpIV & \aMrsgumpV & \aMrsgumpVI & \aMrsgumpVII & \aMrsgumpVIII \tabularnewline
Narrator  & \aNarratorAll & \aNarratorI & \aNarratorII & \aNarratorIII & \aNarratorIV & \aNarratorV & \aNarratorVI & \aNarratorVII & \aNarratorVIII \tabularnewline
\bottomrule
\end{tabular}
\end{table*}

% NLP
% Tiger Treebank Corpus:
% \href{https://www.ims.uni-stuttgart.de}{ims.uni-stuttgart.de}.
To automatically analyze linguistic features of words in their corresponding
sentence, we employed the Python package spaCy v2.2.1
(\href{https://spacy.io}{spacy.io}) employing its accompanying German language model (\texttt{de\_core\_news\_md} trained on the TIGER Treebank corpus \citep{brants2004tiger}.
Non-speech vocalizations were dropped from the sentences before analysis.
We performed analyses regarding part-of-speech (i.e. grammatical tagging or word-category disambiguation), syntactic dependencies, lemmatization, word embedding (i.e. semantic similarity), and if the word is one of the most common words of the German language, i.e. if the word is part of a stop list.
Results of these analyzes were added to the annotation which is available as a Praat TextGrid file (\texttt{annotation/fg\_rscut\_ad\_ger\_speech\_tagged.TextGrid} and as a tab separated file (\texttt{annotation/fg\_rscut\_ad\_ger\_speech\_tagged.tsv}) in accordance with the brain imaging data structure (\href{https://bids.neuroimaging.io/}{BIDS}) \citep{gorgolewski2016bids}).
\todo[inline]{hier bringe ich bereits vor der ``Data availability'' die Pfade/Dateinamen, weil im folgenden die TSV-Datei in der ``Data legend'' beschrieben wird}

\subsection*{Data Legend}
The annotation in TextGrid format comprises 10 tiers: person, sentence (the text of a sentence), words (the text of a word), pos, tag, dep, lemma, stop, vector and phones (pronounciation of a phoneme).
The annotation in TSV format comprises 11 columns: onset, duration, person, text, pos, tag, dep, lemma, stop, vector.
In the following description column header labels are given in parenthesis.

@@

\subsubsection*{Duration (\texttt{end})}
The duration of the sentences, word or phoneme provided the format seconds.milliseconds.

\subsubsection*{Speaker identity (\texttt{person})}
Name of the person that speaks the sentence, word or phoneme (cf. table \ref{tab:speakers} for the ten most often ocurring speakers).

@@

% einige Eigennamen (NE/NN) wurden automatisch durch das Skript korrigiert, weil das Sprachmodell die naheliegenderweise nicht kannte
% eine Liste dieser Eigennamen befindet sich oben in dem entsprechendem Skript
% add_part-of-speech-tagging2textgrid.py
\subsubsection*{Simple part-of-speech tag (\texttt{pos})}
A simple part-of-speech tagging (grammatical tagging; word-category disambiguation) of a word.
The simple part-of-speech tagging tags follow the Universal Dependencies v2 POS tag set (\href{https://universaldependencies.org}{universaldependencies.org}).
See Table \ref{tab:pos} for a description of the labels and the respective counts of all 15 labels.
Additionally in cells of this column, sentences are tagged as "SENTENCE", and phonemes are tagged as "PHONEME" to facilitate sorting and filtering in potential further processing steps of the annotation.

>>>>

% For more detailed article preparation guidelines, please see:
% http://f1000research.com/author-guidelines
\documentclass[10pt,a4paper,onecolumn]{article}
\usepackage{f1000_styles}
\usepackage{units}
\geometry{right=4cm}
% enable the 'endfloat' package to sort all tables and figures to the
% end of the document without having to edit the manuscript
%\usepackage{endfloat}
\usepackage{booktabs}
\usepackage[
	colorlinks=true,
	urlcolor=blue,
	linkcolor=green
]{hyperref}
%% Default: numerical citations
\usepackage[numbers]{natbib}

@@

\author[1, 2]{Christian~O.~Häusler \\ email \href{mailto:c.haeusler@fz-juelich.de}{c.haeusler@fz-juelich.de} }
\author[1, 2]{Michael~Hanke \\ email \href{mailto:m.hanke@fz-juelich.de}{m.hanke@fz-juelich.de} }

@@

\todo{Wort zur Validierungsstudie; Es sind erst 110 von max 300 Worten verbraucht; Werte würde ich eher runden }
\thispagestyle{fancy}
\begin{abstract}
% up to 300 words
Here we present an annotation of speech in the audio-visual movie ``Forrest Gump'' and its audio-description (i.e. the audio-only variant) as an addition to a large public functional brain imaging dataset (\href{www.studyforrest.org}{studyforrest.org}).
The annotation provides information about the exact timing of each of the \aSentencesAll\ spoken sentences, \aWordsAll\ words (including \aPosNonspeechAll\ non-speech vocalizations), \aPhonesAll\ phonemes and its corresponding speaker.
Additionally for every word, we provide lemmatization, two different part-of-speech taggings (simple and detailed), syntactic dependencis within each sentence, and a 300-dimensional semantic representation (word vector).
This annotation facilitates further studies of language perception and semantic processing under conditions of real-life complexity using the \textit{studyforrest} dataset.
\end{abstract}

@@

\section*{Introduction}
\todo[inline]{Bitte am Ende (?) der Einleitung etwas zu "reproducible research" \& "datalad dataset" bringen}
Cognitive neuroimaging research is moving towards studying brain behavior
under conditions of real-life-like complexity. Motion pictures \citep{hasson2008neurocinematics} and continuous narratives \citep{honey2012not, lerner2011topographic} are increasingly utilized as so called ``naturalistic stimuli''.

Because naturalistic stimuli are multi-dimensional and their temporal structure
is usually unknown, data-driven methods like intersubject correlation (ISC)\citep{hasson2004intersubject} or independent component analysis
(ICA)\citep{bartels2004chronoarchitecture} are often used to analyze
corresponding fMRI data.
Nevertheless, model-driven methods like a general linear model (GLM) are necessary to not just explain how but also when and why a brain region is responding to a stimulus \citep{hamilton2018revolution}.
Studies using GLMs based on annotations of a stimulus' temporal structure have
elucidated how the brain responds to visual features of a movie
\citep{bartels2004mapping} or speech-related features of a narrative
\citep{rocca2019language}.

\todo[inline]{In der Tabelle ausführlicher beschrieben als hier (ist okay so, denke ich): ``sentences'' sind hier nicht vollständige Sätze im linguistischen Sinne (Subjekt, Prädikat), sondern auch Ausrufe ("Forrest!") oder mitunter isolierte non-speech "Sätze" ("oouh")}
For this publication, we annotated the onsets and offsets, and content of sentences, single words, and phonemes spoken in the audio-visual movie ``Forrest Gump'' \citep{ForrestGumpMovie} and its audio-description (i.e. the audio-only variant of the movie)\citep{ForrestGumpDVD}(cf. table \ref{tab:overview} for an overview of the annotation's content).
fMRI data of participants watching the audio-visual movie \citep{hanke2016simultaneous} and listening to the audio-description \citep{hanke2014audiomovie} are the core data of the publicly available \textit{studyforrest} dataset (\href{www.studyforrest.org}{studyforrest.org}).
Additional fMRI data of this dataset include the perception of musical genres \citep{hanke2015musical}, and retinotopic mappings and localization of higher visual areas of the same participants \citep{sengupta2016extension}.

@@

\subsection*{Annotation procedure}
\todo[inline]{wo kommen die "preliminaries transcriptions" etc.? Habe die von Medusa; die "most basic" Datei, die das Dataset enthält ist wie abgesprochen der Merge aus Dialog (ungetrennte Sentences), Narrator \& Non-Speech-Annotation = \texttt{annotation/speech-vocal.csv}}
Preliminary, orthographic transcriptions of dialogues (\texttt{german\_dialog.csv}), non-speech vocalizations (e.g. laughing or groaning; \texttt{non\_speech\_vocalization.json}), and the audio-description's narrator
(\texttt{german\_audio\_description.csv}) were merged (\texttt{annotation/speech-vocal.csv}) and converted to Praat's \citep{boersma2019praat} TextGrid format  (\texttt{annotation/speech-vocal.TextGrid}).

% \href{www.fon.hum.uva.nl/praat}{fon.hum.uva.nl/praat}.
The merged transcription containing a rough onset and offset for usually a couple of sentences was edited in Praat v6.0.22 for manual validation against the original audio.
The following steps were done in several passes to iteratively improve the quality of the data:
approximate temporal onsets and offsets were corrected, and intervals containing several sentences were split into intervals containing only one sentence.
When two or more persons were speaking simultaneously the less dominant voice was dropped.
Low volume background speech (especially occurring during music or continuous environmental noise) or low volume non-speech vocalizations that are difficult to recognize by the audience were dropped, too.
\todo[inline]{what is a regular audience? without neurological disorders ;-); Simpel: eine audience, die den Film nicht 45x auf voller Lautstärke über Kopfhörer mit Transkription danebenliegend gehört hat}

% Alignment preparations
% http://prosodylab.org
% https://raw.githubusercontent.com/prosodylab/prosodylab.dictionaries/master/de.dict
We used the Montreal Forced Aligner v1.0.1 \citep{mcauliffe2017montreal} to
algorithmically identify the exact onset of offset of the sentences, and the onset and offset of each word and phoneme embedded in the transcribed sentences.
To enable the aligner to look up the phonemes embedded within each word, we chose a German pronunciation dictionary provided by Prosodylab \citep {gorman2011prosodylab} that uses the Prosodylab PhoneSet to describe the pronounciation of phonemes.
To improve the detection rate of the automatic alignment, the dictionary was manually updated with German words that occur in the stimuli but were originally missing in the dictionary.
% \href{http://mlmlab.org/mfa/dictionaries/english.dict}{mlmlab.org/mfa/dictionaries/english.dict}
The pronunciation of English words and phonemes occurring in the otherwise German audio track was taken from an English pronunciation dictionary (following the ARPAbet PhoneSet) provided by the Montreal Language Modeling lab (\href{http://mlmlab.org/mfa/dictionaries/english.dict}{mlmlab.org}).

% Alignment
\todo[inline]{Nicht erwähnt: Wenn die Korrektur der Phoneme unmöglich war, habe ich einfach die Event-Borders zwischen den Phonemen herausgenommen. Wortbeginn und Wortende stimmen dann mit einem neuen, "künstlich-großen Phoneme" überein.
Das "Groß-Phonem" enthält dann ohne Grenzen zwischen den tatsächlichen Phonemen des Wortes alle Phoneme des Wortes ("es ist kompliziert zu erklären, aber im Grunde ganz einfach). kommt jedoch so selten vor, ist im Grunde "noise" und ist in der Anno sofort als solches zu erkennen.}
The audio track was converted from FLAC to WAV via FFmpeg v4.1.4 \citep{ffmpeg} to meet the aligner's requirements.
This WAV file, the transcription of sentences, and the updated dictionary were
submitted to the aligner that first trained an acoustic model on the data and
then performed the alignment.
The resulting .TextGrid file that contains onsets and offsets of words and phonemes was corrected manually and iteratively in several passes:
In a first step, onsets and offsets on which the automatic alignment performed moderately were corrected in Praat.
Some low volume sentences that are spoken in continuously noisy settings (pursuit, battle, hurricane, rally, disco) were removed due to poor overall alignment performance.
In a second step, the complete sentences of the orthographic transcription were copied into the annotation created by the aligner.
In a third step, the a speaker's identity was added for each sentence (cf. Table \ref{tab:speakers} for the most often occurring speakers).
During every step results from previous steps were checked again for errors.

\begin{table*}[tbp]
\caption{Overview of the annotation's content for the whole stimulus and its individual segments used during fMRI scanning. The label ``sentences'' comprises complete grammatical sentences which are additionally marked in the annotation with a full stop at the end (``my feet hurt.'').
It also comprises questions (``do you want a chocolate?''), exlamations (``run away!), or non-speech vocalizations in quick succession (``ha, ha, ha''), or in isolation (e.g. ``Forrest?'', ``Forrest!'', ``ha'') at time points when speakers switch rapidly. The label ``words'' comprises each word or non-speech vocalization (N=\aPosNonspeechAll) in isolation.}
\label{tab:overview}
\begin{tabular}{llllllllll}
\toprule
\textbf{label} & \textbf{all} & \textbf{1} & \textbf{2} & \textbf{3} & \textbf{4} & \textbf{5} & \textbf{6} & \textbf{7} & \textbf{8}\\
\midrule
sentences  & \aSentencesAll & \aSentencesI & \aSentencesII & \aSentencesIII & \aSentencesIV & \aSentencesV & \aSentencesVI & \aSentencesVII & \aSentencesVIII \tabularnewline
words  & \aWordsAll & \aWordsI & \aWordsII & \aWordsIII & \aWordsIV & \aWordsV & \aWordsVI & \aWordsVII & \aWordsVIII \tabularnewline
phonemes  & \aPhonesAll & \aPhonesI & \aPhonesII & \aPhonesIII & \aPhonesIV & \aPhonesV & \aPhonesVI & \aPhonesVII & \aPhonesVIII \tabularnewline
\bottomrule
\end{tabular}
\end{table*}

\begin{table*}[btp]
\caption{Sentences spoken by the ten most often occurring speakers sorted alphabetically. The narrator only occurrs in the audio-description of the movie.
Overall 97 persons were identified.
Names are mostly identical to the names used in \citep{labs2015portrayed}.}
\label{tab:speakers}
\begin{tabular}{llllllllll}
\toprule
\textbf{name} & \textbf{all} & \textbf{1} & \textbf{2} & \textbf{3} & \textbf{4} & \textbf{5} & \textbf{6} & \textbf{7} & \textbf{8}\\
\midrule
Bubba  & \aBubbaAll & \aBubbaI & \aBubbaII & \aBubbaIII & \aBubbaIV & \aBubbaV & \aBubbaVI & \aBubbaVII & \aBubbaVIII \tabularnewline
Forrest  & \aForrestAll & \aForrestI & \aForrestII & \aForrestIII & \aForrestIV & \aForrestV & \aForrestVI & \aForrestVII & \aForrestVIII \tabularnewline
Forrest (child)  & \aForrestchildAll & \aForrestchildI & \aForrestchildII & \aForrestchildIII & \aForrestchildIV & \aForrestchildV & \aForrestchildVI & \aForrestchildVII & \aForrestchildVIII \tabularnewline
Forrest (v.o.)  & \aForrestvoAll & \aForrestvoI & \aForrestvoII & \aForrestvoIII & \aForrestvoIV & \aForrestvoV & \aForrestvoVI & \aForrestvoVII & \aForrestvoVIII \tabularnewline
Hancock  & \aHancockAll & \aHancockI & \aHancockII & \aHancockIII & \aHancockIV & \aHancockV & \aHancockVI & \aHancockVII & \aHancockVIII \tabularnewline
Jenny  & \aJennyAll & \aJennyI & \aJennyII & \aJennyIII & \aJennyIV & \aJennyV & \aJennyVI & \aJennyVII & \aJennyVIII \tabularnewline
Jenny (child)  & \aJennychildAll & \aJennychildI & \aJennychildII & \aJennychildIII & \aJennychildIV & \aJennychildV & \aJennychildVI & \aJennychildVII & \aJennychildVIII \tabularnewline
Lt. Dan  & \aLtdanAll & \aLtdanI & \aLtdanII & \aLtdanIII & \aLtdanIV & \aLtdanV & \aLtdanVI & \aLtdanVII & \aLtdanVIII \tabularnewline
Mrs. Gump  & \aMrsgumpAll & \aMrsgumpI & \aMrsgumpII & \aMrsgumpIII & \aMrsgumpIV & \aMrsgumpV & \aMrsgumpVI & \aMrsgumpVII & \aMrsgumpVIII \tabularnewline
Narrator  & \aNarratorAll & \aNarratorI & \aNarratorII & \aNarratorIII & \aNarratorIV & \aNarratorV & \aNarratorVI & \aNarratorVII & \aNarratorVIII \tabularnewline
\bottomrule
\end{tabular}
\end{table*}

% NLP
% Tiger Treebank Corpus:
% \href{https://www.ims.uni-stuttgart.de}{ims.uni-stuttgart.de}.
To automatically analyze linguistic features of each words in their corresponding
sentence, we employed the Python package spaCy v2.2.1
(\href{https://spacy.io}{spacy.io}) and its accompanying German language model (\texttt{de\_core\_news\_md} that was trained on the TIGER Treebank corpus \citep{brants2004tiger}.
Non-speech vocalizations were dropped from the sentences before analysis.
We performed analyses regarding part-of-speech (i.e. grammatical tagging or word-category disambiguation), syntactic dependencies, lemmatization, word embedding (i.e. multi-dimensional meaning representation of a word), and if the word is one of the most common words of the German language (i.e. if the word is part of a stop list).
Results of these analyzes were added to the annotation which is available as a (text-based) Praat TextGrid file (\texttt{annotation/fg\_rscut\_ad\_ger\_speech\_tagged.TextGrid} and as a text-based, tab-separated value (TSV) formatted table (\texttt{annotation/fg\_rscut\_ad\_ger\_speech\_tagged.tsv}) in accordance with the brain imaging data structure (\href{https://bids.neuroimaging.io/}{BIDS}) \citep{gorgolewski2016bids}).
\todo[inline]{hier bringe ich bereits vor der ``Data availability'' die Pfade/Dateinamen, weil im Folgenden die TSV-Datei in der ``Data legend'' beschrieben wird}

\subsection*{Data Legend}
The annotation in TextGrid format comprises 10 tiers: person, sentence (the text of a sentence), words (the text of a word), pos, tag, dep, lemma, stop, vector and phones (pronounciation of a phoneme).
The annotation in TSV format comprises 11 columns: onset, duration, person, text (of sentence, word or phoneme), pos, tag, dep, lemma, stop, vector.
In the following description column header labels are given in parenthesis.

@@

\subsubsection*{Duration (\texttt{duration})}
The duration of the sentences, word or phoneme provided the format seconds.milliseconds.

\subsubsection*{Speaker identity (\texttt{person})}
Name of the person that speaks the sentence, word or phoneme. See Table \ref{tab:speakers} for the ten most often ocurring speakers.

@@

\subsubsection*{Simple part-of-speech tag (\texttt{pos})}
A simple part-of-speech tagging (grammatical tagging; word-category disambiguation) of a word.
The simple part-of-speech tagging tags follow the Universal Dependencies v2 POS tag set (\href{https://universaldependencies.org}{universaldependencies.org}).
See Table \ref{tab:pos} for a description of the labels and the respective counts of all 15 labels.
Nouns that spaCy mistook for proper nouns or vice versa were corrected via script.
% s. list at the beginning of script
% code/add_part-of-speech-tagging2textgrid.py
Additionally in cells of this column, sentences are tagged as "SENTENCE", and phonemes are tagged as "PHONEME" to facilitate sorting, and filtering in potential further processing steps.

\begin{table*}[t]
\caption{Simple part-of-speech tagging (pos) performed by the Python package spaCy (\href{https://spacy.io}{spacy.io}).
All 15 labels sorted alpapetically.
Descriptions were taken from spaCy.explain().
Non-speech vocalizations (NONSPEECH) were manually identified.
Counts for the whole stimulus (all) and for each of the eight stimulus segments refer to the audio-description.}
\label{tab:pos}
\begin{tabular}{lllllllllll}
\toprule
textbf{label} & \textbf{description} & \textbf{all} & \textbf{1} & \textbf{2} & \textbf{3} & \textbf{4} & \textbf{5} & \textbf{6} & \textbf{7} & \textbf{8} \\
\midrule
ADJ & \aPosAdj & \aPosAdjAll & \aPosAdjI & \aPosAdjII & \aPosAdjIII & \aPosAdjIV & \aPosAdjV & \aPosAdjVI & \aPosAdjVII \tabularnewline
ADP & \aPosAdp & \aPosAdpAll & \aPosAdpI & \aPosAdpII & \aPosAdpIII & \aPosAdpIV & \aPosAdpV & \aPosAdpVI & \aPosAdpVII & \aPosAdpVIII \tabularnewline
ADV & \aPosAdv & \aPosAdvAll & \aPosAdvI & \aPosAdvII & \aPosAdvIII & \aPosAdvIV & \aPosAdvV & \aPosAdvVI & \aPosAdvVII & \aPosAdvVIII \tabularnewline
AUX & \aPosAux & \aPosAuxAll & \aPosAuxI & \aPosAuxII & \aPosAuxIII & \aPosAuxIV & \aPosAuxV & \aPosAuxVI & \aPosAuxVII & \aPosAuxVIII \tabularnewline
CONJ & \aPosConj & \aPosConjAll & \aPosConjI & \aPosConjII & \aPosConjIII & \aPosConjIV & \aPosConjV & \aPosConjVI & \aPosConjVII & \aPosConjVIII \tabularnewline
DET & \aPosDet & \aPosDetAll & \aPosDetI & \aPosDetII & \aPosDetIII & \aPosDetIV & \aPosDetV & \aPosDetVI & \aPosDetVII & \aPosDetVIII \tabularnewline
NONSPEECH & non-speech vocalization & \aPosNonspeechAll & \aPosNonspeechI & \aPosNonspeechII & \aPosNonspeechIII & \aPosNonspeechIV & \aPosNonspeechV & \aPosNonspeechVI & \aPosNonspeechVII & \aPosNonspeechVIII \tabularnewline
NOUN & \aPosNoun & \aPosNounAll & \aPosNounI & \aPosNounII & \aPosNounIII & \aPosNounIV & \aPosNounV & \aPosNounVI & \aPosNounVII & \aPosNounVIII \tabularnewline
NUM & \aPosNum & \aPosNumAll & \aPosNumI & \aPosNumII & \aPosNumIII & \aPosNumIV & \aPosNumV & \aPosNumVI & \aPosNumVII & \aPosNumVIII \tabularnewline
PART & \aPosPart & \aPosPartAll & \aPosPartI & \aPosPartII & \aPosPartIII & \aPosPartIV & \aPosPartV & \aPosPartVI & \aPosPartVII & \aPosPartVIII \tabularnewline
PRON & \aPosPron & \aPosPronAll & \aPosPronI & \aPosPronII & \aPosPronIII & \aPosPronIV & \aPosPronV & \aPosPronVI & \aPosPronVII & \aPosPronVIII \tabularnewline
PROPN & \aPosPropn & \aPosPropnAll & \aPosPropnI & \aPosPropnII & \aPosPropnIII & \aPosPropnIV & \aPosPropnV & \aPosPropnVI & \aPosPropnVII & \aPosPropnVIII \tabularnewline
SCONJ & \aPosSconj & \aPosSconjAll & \aPosSconjI & \aPosSconjII & \aPosSconjIII & \aPosSconjIV & \aPosSconjV & \aPosSconjVI & \aPosSconjVII & \aPosSconjVIII \tabularnewline
VERB & \aPosVerb & \aPosVerbAll & \aPosVerbI & \aPosVerbII & \aPosVerbIII & \aPosVerbIV & \aPosVerbV & \aPosVerbVI & \aPosVerbVII & \aPosVerbVIII \tabularnewline
X & \aPosX & \aPosXAll & \aPosXI & \aPosXII & \aPosXIII & \aPosXIV & \aPosXV & \aPosXVI & \aPosXVII & \aPosXVIII \tabularnewline
\bottomrule
\end{tabular}
\end{table*}

\subsubsection*{Detailed part-of-speech tag (\texttt{tag})}
% \href{https://www.ims.uni-stuttgart.de/forschung/ressourcen/lexika/germantagsets}{Stuttgart-Tübingen-Tagset}
A detailed part-of-speech tagging of a word following the TIGER Treebank annotation scheme \citep{brants2004tiger} which is based on the Stuttgart-Tübingen-Tagset \citep{schiller1999stts}.
See Table \ref{tab:tag} for a descripion  of the labels and the respective counts of the 15 most often occuring labels. Nouns that spaCy mistook for proper nouns or vice versa were corrected via script.
% s. list at the beginning of script
% code/add_part-of-speech-tagging2textgrid.py
